Importance of depth refugia for reef resilience and intervention on the Great Barrier Reef under future climate change

### 2025-10-05 paper/paper_draft.qmd
Updates to rest of results section

There were substantial differences in the maintenance of coral cover post 2030 between the
GCMs. Figure [-@fig-gcm-cover-trajectories] illustrates the temporal behaviour of reefs
across clusters (Low, Medium and High). Of the GCMs explored, NorESM2-MM had the lowest
Equilibrium Climate Sensitivity (2.5°C) and reefs exhibit continuing recovery and decline
dynamics up to 2075 under this hypothetical climate future. Unsurprisingly, recovery phases
were predominantly seen in the reefs in the High cover cluster, with Medium and Low
exhibiting lesser probability of such periods of recovery. Generally, recovery and increases
in coral cover were visible in simulated results up to 2045. Under NorESM2-MM, the Far
Northern Management area were found to have the most similar temporal behaviour between Low
and High clusters compared to other Management areas (@stbl-cluster-cid) due to the low
degree of coral cover overall (@fig-gcm-cover-trajectories).

Intermediate levels of coral cover maintenance can be seen in ACCESS-ESM1-5 (ECS value of
3.87°C) when compared to the other GCMs (of higher and lower ECS values). Under these
intermediate conditions (in relative terms), there is potential for reefs in the High
cluster to maintain coral cover levels around 20% to 2050. The highest peak DHW levels occur
under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs time series
occur under ACCESS-CM2 (@fig-methods-dhws). Under conditions represented by high ECS value
GCMs - specifically, ACCESS-CM2 (4.66°C), EC-Earth3-Veg (4.30°C) and GFDL-CM4 (4.1°C) - all
reefs are largely in decline across all management areas from 2055 (@stbl-cluster-cid).

![Time series of coral cover (relative to reef area) categorised within each Management area
(rows), for each GCM analysed (columns) based on temporal behaviour over 2030 and 2060. The
time series were clustered into "Low" (green), "Medium" (orange), and "High" (blue) clusters
indicating their relative median coral cover. Solid lines represent the cluster median time
series and bands represent the 95th percentile confidence
intervals.](../figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
fig-align="center" width="14.82cm"}

Comparing across GCMs, it is common for reefs to switch between clusters, including from Low
to High and vice versa (@sfig-gcm-bioregion-cluster-assignment). Of 2197 reefs, 21% (462) of
reefs have consistent cluster assignment across all GCMs. Of the 462 consistently assigned
reefs, 23% are within the Far Northern Management area, with 1%, 55% and 21% occurring in
the Cairns-Cooktown, Townsville-Whitsunday and Mackay-Capricorn Management areas
respectively (@sfig-gcm-cluster-map). Of the 462 consistently assigned reefs, 34%, 18% and
47% are clustered into Low, Medium and High clusters respectively. High levels of assignment
variability across GCMs are expected as due to spatial differences in each given GCM’s DHW
trajectories.

@@

In contrast to the broad temporal behaviour at the Management area scale, the relationship
between reef characteristics and assigned clusters were found to be similar at the bioregion
scale across GCMs. Across bioregions, reefs in the High cluster not only had greater levels
of coral recovery but displayed persistence to 2050
(@sfig-access-esm1-5-bioregion-timeseries). Qualitative representations of bioregion scale
clustering results across assessed reef characteristics can be found in the supplementary
material, with Partial Dependence Analysis results displayed in the main text.

>>>>

There were substantial differences in the maintenance of coral cover post 2030 between the
GCMs (@fig-gcm-cover-trajectories). Of the GCMs explored, NorESM2-MM had the lowest
ECS (2.5°C) and reefs exhibit continuing recovery and decline
dynamics up to 2075 under this hypothetical climate future. Unsurprisingly, recovery phases
were predominantly seen in the reefs in the High cover cluster, with Medium and Low
exhibiting lesser probability of such periods of recovery. Generally, recovery and increases
in coral cover were visible in simulated results up to 2045. Under NorESM2-MM, the Far
Northern Management area was found to have the most similar temporal behaviour between Low
and High clusters compared to other Management areas (@stbl-cluster-cid) due to the low
degree of coral cover overall (@fig-gcm-cover-trajectories).

Intermediate levels of coral cover maintenance can be seen in ACCESS-ESM1-5 (ECS value of
3.87°C) when compared to the other GCMs (of higher and lower ECS values). Under these
intermediate conditions (in relative terms), there is potential for reefs in the High
cluster to maintain coral cover levels around 20% to 2050. The highest peak DHW levels occur
under GFDL-CM4 between 2045 and 2050, while greater increases in overall DHWs time series
occur under ACCESS-CM2 (@fig-methods-dhws). Under conditions represented by high ECS value
GCMs - specifically, ACCESS-CM2 (4.66°C), EC-Earth3-Veg (4.30°C) and GFDL-CM4 (4.1°C) - all
reefs are largely in decline across all management areas from 2035 (@stbl-cluster-cid).

![Time series of coral cover (relative to reef area) clustered within each Management area
(rows), for each GCM analysed (columns) based on temporal behaviour over 2030 and 2060. Time series were clustered into "Low" (green), "Medium" (orange), and "High" (blue) clusters
indicating their relative median coral cover. Solid lines represent cluster median time
series and bands represent 95th percentile confidence
intervals.](../figs/GCM_cover_timeseries_plot.png){#fig-gcm-cover-trajectories
fig-align="center" width="14.82cm"}

Comparing across GCMs, it is common for reefs to switch between clusters, including from Low
to High and vice versa (@sfig-gcm-bioregion-cluster-assignment). Of 2197 reefs, 21% (462) of
reefs have consistent cluster assignment across all GCMs. Of the 462 consistently assigned
reefs, 23% are within the Far Northern Management area, with 1%, 55% and 21% occurring in
Cairns-Cooktown, Townsville-Whitsunday and Mackay-Capricorn Management areas
respectively (@sfig-gcm-cluster-map). Of the 462 consistently assigned reefs, 34%, 18% and
47% are clustered into Low, Medium and High clusters respectively. High levels of assignment
variability across GCMs are expected as due to spatial differences in each given GCM’s DHW
trajectories.

@@

Relationships between reef characteristics and bioregion-scale time series clusters were found to be similar across GCMs. Across bioregions, reefs in the High cluster not only had greater levels
of coral recovery but displayed persistence to 2050
(@sfig-access-esm1-5-bioregion-timeseries). Qualitative representations of bioregion scale
clustering results across assessed reef characteristics can be found in the supplementary
material, with Partial Dependence Analysis results displayed in the main text.
